---
title: "Getting Started with R chapter 3 - Camera trap data examples"
output: html_notebook
---

Prep your R workspace by writing code that will 1) clear R's brain, 2) turn on dplyr and 3) turn on ggplot:

```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
```
Load the data set called "masterDF_test.csv" and call the data frame "DF"

```{r}
library(readr)
DF <- read_csv("~/R/Datasets/masterDF_test.csv")
```
Review from chapter 2: Examine the dataset using the `str` function 
```{r}
str(DF)
```
Let's add a new variable to our data set, Class, which identifies the Class to which each species belongs. It will be a categorical variable, with 2 levels: Mammalia for all the mammals and Aves for the birds.  


1. Use `which` to determine the rows in DF that contain birds. *Note*: If  you are trying to select from a vector of categorical values (e.g. a list of all the birds) you cannot use `==` but must instead use `%in%` to make the selection.
```{r}
Birds<- c("GROUSERUFFED", "BLUEJAY", "OTHERBIRD", "TURKEY", "CROWORRAVEN", "OWL", "GOOSEWILD")
Aves<- which(DF$bin %in% Birds)
```
Now assign those rows to the Class Aves, creating a new column in the dataframe along the way.
```{r}
DF<- DF %>%
 mutate(Class = "Mammalia")

DF$Class[(Aves)]<- "Aves"
#DF$Class<- as.factor(DF$Class)

N<- which(DF$bin == "NOTHINGHERE")
DF$Class[(N)] <- "NA"
```
I had a lot of trouble trying to figure out how to do this code (was overthinking it maybe?)
Other ways to do it?

Now use `filter` and `group_by` to summarize the number of birds and mammals observed, respectively, by round.

```{r}
BirdByRounds<- DF %>%
  group_by(round)%>%
  filter(Class == "Aves")%>%
  summarize(count=n())

MammalsByRounds<- DF%>%
  group_by(round)%>%
  filter(Class == "Mammalia")%>%
  summarize(count=n())
```

Now create a new data frame that contains only the members of the weasel family (mustelidae) from round 4 of the study

```{r}
unique(DF$bin)
Musts<- c("FISHER", "WEASEL", "MINK", "RIVEROTTER")
Musts<- which(DF$bin %in% Musts)
MustsInRound4<- DF%>%
  slice(Musts)%>%
  filter(round==4)

```

Now create a new data frame that contains only the turkeys from round 4 as well as the deer from rounds 2 or 3
```{r}
T<- ("TURKEY")
T<- which(DF$bin %in% T)
D<- ("DEERWHITETAILED")
D<- which(DF$bin %in% D)
T4<- DF %>%
  slice(T)%>%
  filter(round == 4)
D23<- DF%>%
  slice(D)%>%
  filter(round == 2 | round == 3)
```




